---
title: "Adult"
author: "R Bridge Meetup 3"
date: "July 28, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The Adult dataset in the UCI repository can be used to "predict whether income exceeds $50K/yr based on census data."

Most data science workflows involve loading and subsetting data.  This bare bones code demonstrates a reasonable way to start.

## What this code does
* Load Adult data from UCI repository into an R data frame.  
* Select columns related to age, salary, and education
* Rename the column headers

Alternatively, we could have renamed the columns when we loaded our data, which would have been a little more work (since we rename columns that we subsequently filter out), but would be more readable.

RCurl is a powerful "heavyweight" package for working with web data.  Other approaches include using the read_csv() function in Hadley Wickham's `readr` package.

#Other work that could be done
* "bin" ages into groups
* order education factors:  PreSchool < ... < Doctorate
* how might you visually show the relationships between the 3 variables?

```{r}
#install.packages("RCurl")

require(RCurl)

url <- "https://archive.ics.uci.edu/ml/machine-learning-databases/adult/adult.data"

adult_data <-getURL(url)
adult_df <- read.csv(text=adult_data,header=F,sep=",")

dim(adult_df)
head(adult_df)
str(adult_df)

# include only columns corresponding to age, education, and salary
adult_df2 <- subset(adult_df,select = c(V1,V4,V15))

head(adult_df2)

colnames(adult_df2) <- c("Age","Education","Salary")
head(adult_df2)

```

