---
title: "graphics"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## graphics examples

Based on examples in R Graphics Cookbook

See also: https://www.rstudio.com/wp-content/uploads/2015/12/ggplot2-cheatsheet-2.0.pdf

```{r s, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
#install.packages(c("ggplot2","gcookbook"))
library(ggplot2)
library(gcookbook)

data(package="gcookbook")
data(package="MASS")

vignette(package="ggplot2")
vignette(package="gcookbook")
```

## Creating a Histogram

Use histograms to view the distribution of one-dimensional data

```{r}
str(mtcars)
#View(mtcars)
ggplot(mtcars, aes(x=mpg)) + geom_histogram(binwidth=4)
```

## Creating a Box Plot

box plots compare distributions
pass in: factor of x values, and vector of y values

```{r}

#View(ToothGrowth)

ggplot(ToothGrowth, aes(x=supp, y=len)) + geom_boxplot()

```

## Making a Basic Scatter Plot

Use geom_point(), and map one variable to x and one to y.

default shape: #16 - solid circles
default size: 2


```{r}
str(heightweight)
heightweight[,c("ageYear","heightIn")]

ggplot(heightweight, aes(x=ageYear, y=heightIn)) + geom_point()

```

## Group Data Points by a Variable Using Shape or Color

map the grouping variable to shape or colour
grouping variable must be categorical (factor or character vector)

```{r}
heightweight[,c("sex","ageYear","heightIn")]
ggplot(heightweight, aes(x=ageYear, y=heightIn, shape=sex )) + geom_point()
ggplot(heightweight, aes(x=ageYear, y=heightIn, colour=sex )) + geom_point()

ggplot(heightweight, aes(x=ageYear, y=heightIn, shape=sex, colour=sex )) + geom_point()

ggplot(heightweight, aes(x=ageYear, y=heightIn, shape=sex, colour=sex )) +
  geom_point() +
  scale_shape_manual(values=c(1,2)) +
  scale_colour_brewer(palette="Set1")
```

## facets

```{r}
p <- ggplot(heightweight, aes(x=ageYear, y=heightIn )) + geom_point()

p + facet_grid( sex ~ .)

```

## mpg dataset (included with ggplot2 package)
```{r}
dim(mpg)
str(mpg)
#View(mpg)
```

## mtcars dataset (included with base R datasets)
https://stat.ethz.ch/R-manual/R-devel/library/datasets/html/mtcars.html

```{r}
dim(mtcars)
str(mtcars)
#View(mtcars)
```

## diamonds dataset (included with ggplot2 package)
```{r}
dim(diamonds)
str(diamonds)
#View(diamonds)
```

## Making a Cleveland dot plot

See also: https://www.perceptualedge.com/articles/b-eye/dot_plots.pdf

```{r}
tophit <- tophitters2001[1:25,c("name","lg","avg")]
ggplot(tophit,aes(x=avg,y=name)) + geom_point()

ggplot(tophit,aes(x=avg,y=reorder(name,avg))) + geom_point()

ggplot(tophit,aes(x=avg,y=reorder(name,avg))) + geom_point(size=3) +
  theme_bw() +
  theme(panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.major.y = element_line(colour="grey60", linetype = "dashed"))

ggplot(tophit,aes(x=reorder(name,avg), y=avg)) + geom_point(size=3) +
  theme_bw() +
  theme(axis.text.x = element_text(angle=60, hjust=1),
        panel.grid.major.y = element_blank(),
        panel.grid.minor.y = element_blank(),
        panel.grid.major.x = element_line(colour="grey60", linetype = "dashed"))

nameorder <- tophit$name[order(tophit$lg, tophit$avg)]

# turn name into a factor, with levels in the order of nameorder
tophit$name <- factor(tophit$name, levels=nameorder)

ggplot(tophit, aes(x=avg, y=name)) +
    geom_segment(aes(yend=name), xend=0, colour="grey50") +
    geom_point(size=3, aes(colour=lg)) +
    scale_colour_brewer(palette="Set1", limits=c("NL","AL")) +
    theme_bw() +
    theme(panel.grid.major.y = element_blank(),   # No horizontal grid lines
          legend.position=c(1, 0.55),             # Put legend inside plot area
          legend.justification=c(1, 0.5))

# using facets
ggplot(tophit, aes(x=avg, y=name)) +
    geom_segment(aes(yend=name), xend=0, colour="grey50") +
    geom_point(size=3, aes(colour=lg)) +
    scale_colour_brewer(palette="Set1", limits=c("NL","AL"), guide=FALSE) +
    theme_bw() +
    theme(panel.grid.major.y = element_blank()) +
    facet_grid(lg ~ ., scales="free_y", space="free_y")
  
```
