---
title: "R Bridge Final Project - Gilde"
author: "Kyle Gilde"
date: "August 6, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load & Re-shape the Data

```{r cars echo=F}
require(ggplot2)
require(ggthemes)
require(plyr)
require(hflights)
my_hflights <- hflights

my_hflights$DayOfWeek[my_hflights$DayOfWeek == 1] <- "1 Sunday"
my_hflights$DayOfWeek[my_hflights$DayOfWeek == 2] <- "2 Monday"
my_hflights$DayOfWeek[my_hflights$DayOfWeek == 3] <- "3 Tuesday"
my_hflights$DayOfWeek[my_hflights$DayOfWeek == 4] <- "4 Wednesday"
my_hflights$DayOfWeek[my_hflights$DayOfWeek == 5] <- "5 Thursday"
my_hflights$DayOfWeek[my_hflights$DayOfWeek == 6] <- "6 Friday"
my_hflights$DayOfWeek[my_hflights$DayOfWeek == 7] <- "7 Saturday"

my_hflights$DayOfWeekType <- ifelse(my_hflights$DayOfWeek == "1 Sunday" | my_hflights$DayOfWeek == "7 Saturday", "Weekend", "Weekday")

my_hflights$Date <- as.Date(ISOdate(my_hflights$Year, my_hflights$Month, my_hflights$DayofMonth))
my_hflights <- my_hflights[ , !(names(my_hflights) %in% c("Year", "Month", "DayofMonth"))]

my_hflights$DepTime <- my_hflights$DepTime/100


my_hflights <- my_hflights[c(20 , 19,  1:18)]




str(my_hflights)
#a.my_hflights <- my_hflights[c(20, 13)]


```

## Question 1: Were there more flights on certain days of the week?

```{r Question 1, echo=FALSE}

countflights <- table(my_hflights$DayOfWeek)
barplot(countflights, las = 2, col = "blue",  main = "Number of Flights by Day")
```

I expected there to be more flights on Friday and Saturday, but the histogram shows that there were more flights on Wednesdays, Thursdays and Sundays. 

## Question 2: Which carriers had the greatest average Departure Delays (sum of DepDelay/# of flights)?

```{r Question 2}
sub_hflights <- na.omit(subset(my_hflights, Cancelled + Diverted == 0, select = c(DepDelay, TaxiIn)))

sum_DepDelay <- aggregate(DepDelay ~ UniqueCarrier, my_hflights, sum)
ct_flights <- aggregate(FlightNum ~ UniqueCarrier, my_hflights, length)
merge_df <- merge(x = sum_DepDelay, y = ct_flights, by.x = "UniqueCarrier", by.y = "UniqueCarrier")
merge_df$avg_DepDelay <- merge_df$DepDelay / merge_df$FlightNum
my_output <- merge_df[order(-merge_df$avg_DepDelay),]
names(my_output) <- c("Carrier", "Sum of Departure Delay (Minutes)", "Count of Flights", "Avg Departure Delay (Minutes)")
my_output



```
I used the aggregate, merge and order functions to create this table that shows the WN carrier as having greatest average departure delay at more than 13 minutes.

## Question 3: 

```{r Question 3, echo=F}

#My Notes
#counts NAs in 1 column and all columns
sum(is.na(df$x))
na_count <-sapply(x, function(y) sum(length(which(is.na(y)))))

# Change Characters and Vectors in DFs 
my_hflights <- lapply(my_hflights, as.character)
my_hflights <- lapply(my_hflights, factor)

ggplot(na.omit(my_hflights), aes(x=DepDelay, y=Distance)) + geom_point()



```



