---
title: "R Bridge Final Project - Gilde"
author: "Kyle Gilde"
date: "August 6, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load, Re-shape & Print Sample of the Data Set

```{r Load, echo = FALSE}
require(ggplot2)
require(ggthemes)
require(plyr)
require(hflights)
my_hflights <- hflights

#Column name changes
names(my_hflights)[names(my_hflights)=="UniqueCarrier"] <- "Carrier"
names(my_hflights)[names(my_hflights)=="DepTime"] <- "DepartureTime"
names(my_hflights)[names(my_hflights)=="DepDelay"] <- "DepartureDelay"
names(my_hflights)[names(my_hflights)=="FlightNum"] <- "FlightNumber"
  
#Change DepartureTime to uniform string  
my_hflights$DepartureTime <- as.character(my_hflights$DepartureTime)
my_hflights$DepartureTime <- ifelse(nchar(my_hflights$DepartureTime) == 3, paste("0", my_hflights$DepartureTime, sep = ""), my_hflights$DepartureTime)
my_hflights$DepartureTime <- paste(substr(my_hflights$DepartureTime, 1, 2), ":", substr(my_hflights$DepartureTime, 3, 4), sep = "")

unique(my_hflights$DepartureTime)

#Distinguish between weekend and weekdays
my_hflights$DayOfWeekType <- ifelse(my_hflights$DayOfWeek %in% c(1, 7), "Weekend", "Weekday")

#add Date column
my_hflights$Date <- as.Date(ISOdate(my_hflights$Year, my_hflights$Month, my_hflights$DayofMonth))

#Select specific columns & Print a head sample
my_hflights <- my_hflights[c(23 , 4, 22, 5, 7, 8, 13, 19, 21)]

knitr::kable(head(my_hflights, n = 10), row.names = FALSE)
#str(my_hflights)
```

## Question 1: Were there more flights on certain days of the week?

```{r Question 1, echo = FALSE}

countflights <- table(my_hflights$DayOfWeek)
barplot(countflights, names.arg = c("Sunday", "Monday", "Tuesday", "wednesday", "Thurday", "Friday", "Saturday"),las = 2, col = "blue",  main = "Number of Flights by Day", cex.names=.9)
```

### Results: I expected there to be more flights on Friday and Saturday, but the barplot shows that there were more flights on Wednesdays, Thursdays and Sundays. 

## Question 2: Which carrier had the greatest average Departure Delay (sum of DepartureDelay / # of flights)?

```{r Question 2, echo = FALSE}
sub_hflights <- na.omit(subset(my_hflights, Cancelled + Diverted == 0, select = c(Carrier, DepartureDelay, FlightNumber)))

sum_DepartureDelay <- aggregate(DepartureDelay ~ Carrier, sub_hflights, sum)

ct_flights <- aggregate(FlightNumber ~ Carrier, sub_hflights, length)

merge_df <- merge(x = sum_DepartureDelay, y = ct_flights, by.x = "Carrier", by.y = "Carrier")


merge_df$avg_DepartureDelay <- round(merge_df$DepartureDelay / merge_df$FlightNumber, digits = 2)
my_output <- merge_df[order(-merge_df$avg_DepartureDelay),]

names(my_output) <- c("Carrier", "Sum of Departure Delays (Minutes)", "Count of Flights", "Avg Departure Delay (Minutes)")
knitr::kable(my_output, row.names = FALSE)



```

Avg Departure Delay (Minutes) = Sum of Departure Delays / Count of Flights

### Results: I used the aggregate, merge and order functions to create this table that shows the WN carrier had the greatest average departure delay at nearly 13.5 minutes.

## Question 3: Do departure delays accummulate more at certain times of the day?

```{r Question 3, echo=FALSE}
my_agg <- aggregate(DepartureDelay ~ DepartureTime + Carrier, my_hflights, sum)

ggplot(my_agg, aes(x = DepartureTime, y = DepartureDelay)) + geom_point(aes(color = Carrier))  + labs(title = "Departure Times & Departure Delays", x = "Departure Time", y = "Sum of Departure Delay Minutes") 
        + scale_x_continuous(breaks = TRUE) + theme_economist()

#My Notes
#counts NAs in 1 column and all columns
# sum(is.na(df$x))
# na_count <-sapply(x, function(y) sum(length(which(is.na(y)))))
# 
# # Change Characters and Vectors in DFs 
# my_hflights <- lapply(my_hflights, as.character)
# my_hflights <- lapply(my_hflights, factor)
# 
# ggplot(na.omit(my_hflights), aes(x=DepartureDelay, y=Distance)) + geom_point()

```

### Results: Yes, the departure delay minutes appear to increase from the morning to later at night.



# ...The End!
